\documentclass{article}
\usepackage{graphicx} % Required for inserting images

\title{Digital Control Systems Coursework: Omar Ben-Gacem}
\author{Omar Ben-Gacem}
\date{January 2025}



%%%%%%%%%%%%     Formatting     %%%%%%%%%%%%

\usepackage[sorting=none]{biblatex} %Imports biblatex package

\addbibresource{bibliography.bib} %Import the bibliography file
\usepackage[margin=1in,headsep=.5in]{geometry}
\usepackage{amsmath} % for bmatrix and pmatrix
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{caption}
\usepackage{fancyhdr}
\usepackage{subcaption}
\usepackage{gensymb}
\usepackage{float}
\usepackage{ragged2e}
\usepackage{booktabs} % For better table formatting
\usepackage[export]{adjustbox}
\geometry{
 a4paper,
 total={170mm,257mm},
 left=19mm,
 right=19mm,
 top=23mm,
 bottom=23mm
 }
\renewcommand{\sectionmark}[1]{%
    \markright{\MakeUppercase{#1}}%
}



\title{Digital Control Systems Coursework Submission}
\author{Omar Ben-Gacem}
\pagestyle{fancy}
\fancyhead[RO]{{\rightmark}}
\date{November 2024}
\lhead{Omar Ben-Gacem\\01883771\\Digital Control Systems Coursework}






%%%%%%%%%%%%     Formatting     %%%%%%%%%%%%


\begin{document}


\section*{Executive Summary}
\input{equations/reachability}


\section{System Dynamics (Part A)}
The system being modeled is an inverted pendulum attached to a sliding cart. The target is to control the liner displacement of the card in order to hold the pendulum in an inverted position. Equation \ref{eqofm} gives the Newtonian equations of motion for the cart-pendulum system.

\input{equations/equations_of_motion}

\subsection{A1) Standard Form}
To write the cart-pendulum system in standard form, the state variable $\underline{x}$ and the input $\underline{u}$ must be identified. The applied force $\mu(t)$ acts as the only input to this system. By inspection of the equations of motion, it is trivial to show that displacement $s$ and angle $\phi$ both have first- and second-order differential terms, thus making state variable $\underline{x} = [s, \dot s, \phi, \dot\phi]^T$. Equation \ref{eqofm} can be rearranged and reduced to give Equation \ref{fofx} which acts as the state update function.

\input{equations/f_of_x}

\subsection{A2) Free Response Equilibrium}
Setting $u=\mu(t)=0$ gives the free response of the system, where the equilibrium of this system are values of $\underline{\dot x}=\underline{0}$. The values of $\underline{x}$ that satisfy this criteria are shown below in Equation \ref{ueqzero}.

\input{equations/ueqzero}

Equation \ref{fofx} shows that the displacement of the cart, $s$, has no impact on the systems ability to reach equilibrium. Thus the possible values of $s$ is determined to be the set of real numbers. In practice, it is bounded to be a reasonable amount of translational motion. The values of $\phi$ is set to be any period of $\pi$, meaning the pendulum is either vertically up or down. 

\subsection{A3) Linearization About an Equilibrium}
Given a point to linearize about, the state space model of the non-linear system $f(\underline{x},\underline{u})$ can be derived using the Jacobian operator to give the Jacobian matrix. The linearization should yeild an equation of the form $\dot x = Ax+Bu$. Equation \ref{jacobian} shows the taking of the Jacobian vector at the point $\underline{x}=\underline{0}$, to get the matrix $A$.

\input{equations/jacobian}

To find matrix $B$, the Jacobian is taken again, however, this time with respect to input $u$. Equation \ref{jB} shows the derivation of the input Jacobian for equilibrium $\underline{x}=\underline{0}$.

\input{equations/jacobian_b}


This can be put into full state space representation as shown in Equation \ref{linized}. This gives the full state-space representation of the cart-pendulum system, and can be used to describe the local behavior of the system. about the point the equilibrium was taken.
\input{equations/full_ss}

\subsection{A4) Linear Dynamics}
The system dynamics

\subsection{A5) Reachability}

\subsection{A6) Observability}

\end{document}
